---
title: "Data Visualization"
format: html
editor: visual
---

## Laying down the foundation

Before we go in to making our maps, we must first download all of the necessary packages in order to make the maps we desire. I made sure to bring the tmap and mapview packages over just to make sure I had multiple options for my map visualization.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tigris)
library(sf)
library(tmap)
library(tmaptools)
library(htmltools)
library(janitor)
library(rmapshaper)
library(here)
library(tidycensus)
library(mapview)
library(leafsync)
library(leaflet.extras2)
options(tigris_class = "sf")

```

Next, I made sure to input the census API key in order to access the census data that I ultimately used to find the variables for my mapping, age and income.

```{r}
# The census key for MapView mapping

census_api_key("2a6f8c21a30d3024e038d67d7d4eba647dc79cd4", overwrite=TRUE)

```

## Creating an Interactive County Census Map Visualization using MapView

I ultimately landed on using MapView in order to create a visually appealing map visualization that any consumer may be able to understand. To create the map, I followed the following steps and code in order to produce a map of Virginia demonstrating the median age and another comparing the median income of different counties in order to find any relation between age and income in Virginia.

```{r}
# Establish the potential variables I want to use in the analysis of median age in Virginia counties.

myvars <- c(totalpop = "B01003_001",
            medincome = "B19013_001",
            medage = "B01002_001")
# Make a table with all of the variables chosen 

va_counties<- get_acs(geography = "county",
                       variables = c(myvars),
                       state = "VA",
                       output = "wide",
                       geometry = TRUE)
# Clean up the column names by taking out the "E"

colnames(va_counties) <- sub("E$", "", colnames(va_counties))

# Customize the labels and popups for the map by creating multiple labels for each map and a common-use popup

mylabelone <- glue::glue("{va_counties$NAM}: {va_counties$medage} years old")
mylabeltwo <- glue::glue("{va_counties$NAM}: ${va_counties$medincome}")

mypopup <- glue::glue("<strong>{va_counties$NAM}</strong><br />
                      Total Population: {va_counties$totalpop}<br />
                      Median Age in the County: {va_counties$medage}<br />
                      Median Income in the County: ${va_counties$medincome}") %>% 
  lapply(htmltools::HTML)

# Write the code and put in a color scheme that I want to create the final interactive maps of both age and income, split between blue to demonstrate income and pink to demonstrate age.

vamap_age <- mapview(va_counties, zcol = "medage", 
         col.regions = RColorBrewer::brewer.pal(9, "PuRd"), 
         alpha.regions = 1,
         popup = mypopup,
        label = mylabelone)

vamap_income <- mapview(va_counties, zcol = "medincome", 
         col.regions = RColorBrewer::brewer.pal(9, "Blues"), 
         alpha.regions = 1,
         popup = mypopup,
        label = mylabeltwo)

sync(vamap_income, vamap_age)

vamap_age | vamap_income
```

## Conclusions

Using the map visualization and census data, one can find that the county with the highest median income in Virginia is Northumberland County with a median age of 59.8. The county with the lowest median age was Lexington City County with a median age of 22.4, probably due to the dense college-age population in the area. In income, however, the county with the highest median income was Loudoun County with a median income of 156,821 dollars and the lowest was Dickenson County at 33,905 dollars. Virginia displays an interesting trend with income and age, in that it shows a more bell-curved trend rather than a positive or negative correlation between income and age. In Virginia being my home state, I find this pattern to be quite indicative of the city-sprawling areas in the state, but question the role that older retirees and their incomes have in skewing the results of this data collection.
